Making Sequential Programs Deterministic

### 2019-04-06 experiments.tex
Some edits in related work
Edits in NonDex and DeFlaker experiments

We ran NonDex on the same commit of Randoop as we used for the case study
of \cref{sec:randoop-case-study}.  This version contains all the nondeterminism bugs that
\theDeterminismChecker found.

We modified Randoop by deleting tests that were skipped by its buildfile,
because the NonDex Gradle plugin does not respect those settings.
After that, NonDex ran without problems on Randoop.
NonDex reported no flaky tests.
%\todo{Revise the following text.
%  I renumbered the bugs without realizing that this section refers to them
%  by number, and I don't know which one these refer to any
%  more.
%  Please use cross-references or macros.  Also, this needs to be expanded
%  to account for the additional bugs.}
We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
\bugHashSet and \bugClasspath have corresponding test cases whereas the remainder of the bugs do not.
As NonDex is a \textit{flaky test} detection tool, detecting \bugTimestampOutput and \bugTimestampOutput 
is outside of its scope.
The reason for nondeterminism in \bugClasspath was a call to the \<System.getProperty()> method which
is not modeled by NonDex.

@@

    \begin{subfigure}[b]{0.95\textwidth}
        \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
        for (Entry<String, FieldType> entry : fields.entrySet()) {
            ...
        }
        ...
    }
        \end{verbatim}
        \caption{Annotated source code from ActionGenerator.}
        \label{code-actiongenerator}
    \end{subfigure}    
    \caption{Errors detected by \TheDeterminismChecker in NonDex benchmarks.}
    \label{fig:nondex-source}
\end{figure}

@@

The ActionGenerator code (\cref{code-actiongenerator}) is similar.  Other
code assumes that \<entry> is deterministic, but annotating it as \<@Det>
leads to a warning from \theDeterminismChecker that iterating over
\<fields.entrySet()> (which is itself \<@OrderNonDet>) yields \<@NonDet>
\<Entry> values..

@@

\TheDeterminismChecker is capable of detecting nondetermism in test
cases, as some of these examples are.
We wrote specifications for the
JUnit library (in addition to the JDK and some other libraries).
The specifications are the \<@Det> type qualifier on the formal parameters
to the \<assert>* methods.
\TheDeterminismChecker reports any \<junit> test that passes nondeterministic arguments to
\<assert>*, even if the test case is not executed.

@@

\section{Comparison to Deflaker}
Deflaker~\cite{deflaker}, like NonDex, reports tests that could be flaky. 
We were unable to run DeFlaker on Randoop, because Deflaker works with
projects built with Maven, but Randoop
uses Gradle for its build system. We plan to convert Randoop's build system to Maven and
perform this case study.
%Ideally, we would have 
%liked to run all three tools (NonDex, DeFlaker, and \TheDeterminismChecker) on an open source project
%that wasn't explored by any of the tool for a fair comparison. Unfortunately, there are issues with the
%DeFlaker plugin and we weren't able to run it on any project. We have created a Github issue for the same
%in its repository.
%\todo{Deflaker works now. Tried on retrofit: only works where there is sufficient change in code coverage between head and its parent.
%Looked throught= about 10 commits, didn't find differential code coverage and therefore no flakes}
%\todo{Deflaker requires Maven build. Randoop is on Gradle.}

@@

Four of the seven flaky tests (two in \<achilles>, one each in \<jackrabbit-oak> and \<togglz>)
were caused by a race condition, which \theDeterminismChecker
cannot detect.  (This is a strength of DeFlaker over \theDeterminismChecker.)
We were unable to build \<togglz> and \<nutz> which had one flakiness issue each.
However, we extracted the source code causing the flakiness in these repositories into test cases
after looking at the corresponding issues on Github. The bug in \<togglz> was caused by a copy method that iterated over
an \<OrderNonDet Set> and expected it to be deterministic. \TheDeterminismChecker correctly flagged
an error in the loop that iterated over the \<Set>.
The flakiness in \<nutz> was a result of printing a response received over http.
Since network operations are nondeterministic, we annotated the method in \<nutz> that
returns this response as \<NonDet> which lead \TheDeterminismChecker to report an error at the print statement. 
\TheDeterminismChecker was able to find the source of flakiness in \<checkstyle>. This bug was in a test
case that treated an array returned by \<Class.getDeclaredConstructors()>
as deterministic. This is erroneous because \<getDeclaredConstructors> returns an order nondeterministic array.

>>>>

We ran NonDex on the same commit of Randoop as we used for the case study
of \cref{sec:randoop-case-study}.  This version contains all
\numRandoopBugs nondeterminism bugs that
\theDeterminismChecker found.

We modified Randoop by deleting tests that were skipped by its buildfile,
because the NonDex Gradle plugin does not respect those settings.
After that, NonDex ran without problems on Randoop.
NonDex reported no flaky tests.
%\todo{Revise the following text.
%  I renumbered the bugs without realizing that this section refers to them
%  by number, and I don't know which one these refer to any
%  more.
%  Please use cross-references or macros.  Also, this needs to be expanded
%  to account for the additional bugs.}

We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
\bugHashSet and \bugClasspath have corresponding test cases whereas the remainder of the bugs do not.
As NonDex is a \textit{flaky test} detection tool, detecting \bugTimestampOutput and \bugTimestampOutput 
is outside of its scope.
The reason for nondeterminism in \bugClasspath was a call to the \<System.getProperty()> method which
is not modeled by NonDex.

@@

    \begin{subfigure}[b]{0.95\textwidth}
        \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
        for (Entry<String, FieldType> entry : fields.entrySet()) {
            ...
        }
        ...
    }
        \end{verbatim}
        \caption{Nondeterministic source code from ActionGenerator.}
        \label{code-actiongenerator}
    \end{subfigure}    
    \caption{Errors detected by \TheDeterminismChecker in NonDex benchmarks.}
    \label{fig:nondex-source}
\end{figure}

@@

The ActionGenerator code (\cref{code-actiongenerator}) is similar.  Other
code assumes that \<entry> is deterministic, but annotating it as \<@Det>
leads to a warning from \theDeterminismChecker that iterating over
\<fields.entrySet()> (which is itself \<@OrderNonDet>) yields \<@NonDet>
\<Entry> values.

@@

\TheDeterminismChecker is capable of detecting nondetermism in test
cases, as some of the NonDex examples are.
We wrote specifications for the
JUnit library (in addition to the JDK and some other libraries).
The specifications are the \<@Det> type qualifier on the formal parameters
to the \<assert>* methods.
\TheDeterminismChecker reports any \<junit> test that passes nondeterministic arguments to
\<assert>*, even if the test case is not executed.

@@

\section{Comparison to Deflaker}
Deflaker~\cite{deflaker}, like NonDex, reports tests that could be flaky. 
We were unable to run DeFlaker on Randoop, because Deflaker works with
projects built with Maven, but Randoop
uses Gradle for its build system.
%% Don't make promises: they give reviewers an excuse to reject the paper
%% as not a completed research project.
% We plan to convert Randoop's build system to Maven and perform this case study.
%Ideally, we would have 
%liked to run all three tools (NonDex, DeFlaker, and \TheDeterminismChecker) on an open source project
%that wasn't explored by any of the tool for a fair comparison. Unfortunately, there are issues with the
%DeFlaker plugin and we weren't able to run it on any project. We have created a Github issue for the same
%in its repository.
%\todo{Deflaker works now. Tried on retrofit: only works where there is sufficient change in code coverage between head and its parent.
%Looked throught= about 10 commits, didn't find differential code coverage and therefore no flakes}
%\todo{Deflaker requires Maven build. Randoop is on Gradle.}

@@

Four of the seven flaky tests (two in \<achilles>, one each in \<jackrabbit-oak> and \<togglz>)
were caused by a race condition, which \theDeterminismChecker
cannot detect.  (This is a strength of DeFlaker over \theDeterminismChecker.)
We were unable to build \<togglz> and \<nutz> which had one flakiness issue each.
However, we extracted the source code causing the flakiness in these repositories into test cases
after looking at the corresponding issues on Github. The bug in \<togglz> was caused by a copy method that iterated over
an \<OrderNonDet Set> and expected it to be deterministic. \TheDeterminismChecker correctly flagged
an error in the loop that iterated over the \<Set>.
The flakiness in \<nutz> was a result of printing a response received over http.
Since network operations are nondeterministic, we annotated the method in \<nutz> that
returns this response as \<NonDet> which led \TheDeterminismChecker to report an error at the print statement. 
\TheDeterminismChecker also found the source of flakiness in \<checkstyle>. This bug was in a test
case that treated an array returned by \<Class.getDeclaredConstructors()>
as deterministic. This is erroneous because \<getDeclaredConstructors> returns an order-nondeterministic array.
